The Efficient Handling of Guards \\in the Design of RPython's Tracing JIT

### 2012-09-07 talk/vmil2012/paper.tex
Trying to save some space
safe some lines

\begin{figure}[ht]
\includegraphics[width=0.4\textwidth]{figures/resume_data.pdf}
\caption{The resume data for Figure~\ref{fig:trace-log}}
\label{fig:resume-data}
\end{figure}

@@

SPUR~\cite{bebenita_spur:_2010} is a tracing JIT compiler
for a CIL virtual machine.
It handles guards by always generating code for every one of them
that transfers control back to the unoptimized code.
Since the transfer code needs to reconstruct the stack frames
of the unoptimized code,
the transfer code is quite large.

@@

Gal et. al~\cite{Gal:2006} describe how in the HotpathVM, a JIT for a Java VM, they experimented
with having one generic compensation code block, like the RPython JIT, that
uses a register variable mapping to restore the interpreter state. Later this
was replaced by generating compensation code for each guard which produced a
lower overhead in their benchmarks. HotpathVM also records secondary traces
starting from failing guards that are connected directly to the original trace.
Secondary traces are compiled by first restoring the register allocator state to
the state at the side exit. The information is retrieved from a mapping stored
in the guard that maps machine level registers and stack to Java level stack
and variables.

@@

Deoptimization in method-based JITs is used if one of the assumptions
of the code generated by a JIT-compiler changes.
This is often the case when new code is added to the system,
or when the programmer tries to debug the program.

Deutsch et. al.~\cite{deutsch_efficient_1984} describe the use of stack descriptions
to make it possible to do source-level debugging of JIT-compiled code.
Self uses deoptimization to reach the same goal~\cite{holzle_debugging_1992}.
When a function is to be debugged, the optimized code version is left
and one compiled without inlining and other optimizations is entered.
Self uses scope descriptors to describe the frames
that need to be re-created when leaving the optimized code.
The scope descriptors are between 0.42 and 1.09 times
the size of the generated machine code.
All the information needed for debugging together
is between 1.22 and 2.33 times the size of generated machine code,
according to the paper.

@@

Additionally we have presented experimental data collected using the standard PyPy
benchmark set to evaluate previous observations and assumptions about guards. Our
experiments confirmed that guards are a very common
operation in traces. At the same time guards are associated with a high
overhead, because for all compiled guards information needs to be
stored to restore the execution state in case of a bailout. The measurements
showed that the compression techniques used in PyPy effectively reduce the
overhead of guards, but they still produce a significant overhead. The results
also showed that guard failure is a local event: there are few
guards that fail at all, and even fewer that fail very often.
These numbers validate the design decision of reducing the overhead of
successful guard checks as much as possible while paying a higher price in the
case of bailout due to having to decode a compressed state representation.
The compressed state representation reduces the memory footprint of rarely
used data.

>>>>

\begin{figure}[ht]
\includegraphics[width=0.4\textwidth]{figures/resume_data.pdf}
\vspace{-3mm}
\caption{The resume data for Figure~\ref{fig:trace-log}}
\label{fig:resume-data}
\end{figure}

@@

SPUR~\cite{bebenita_spur:_2010} is a tracing JIT compiler
for a CIL virtual machine.
It handles guards by always generating code for every one of them
that transfers control back to the unoptimized code.
Since the transfer code needs to reconstruct the stack frames
of the unoptimized code,
the transfer code is large.

@@

Gal et. al~\cite{Gal:2006} describe the HotpathVM, a JIT for a Java VM. They experimented
with having one generic compensation code block, like the RPython JIT, that
uses a register variable mapping to restore the interpreter state. Later this
was replaced by generating compensation code for each guard which produced a
lower overhead in their benchmarks. HotpathVM also records secondary traces
starting from failing guards that are connected directly to the original trace.
Secondary traces are compiled by first restoring the register allocator state to
the state at the side exit. The information is a mapping stored
in the guard between machine level registers and stack to Java level stack
and variables.

@@

Deoptimization in method-based JITs is used if one of the assumptions
of the code generated by a JIT changes.
This is often the case when new code is added to the system,
or when the programmer tries to debug the program.

Deutsch et. al.~\cite{deutsch_efficient_1984} use stack descriptions
to make it possible to do source-level debugging of JIT-compiled code.
Self uses deoptimization to reach the same goal~\cite{holzle_debugging_1992}.
When a function is to be debugged, the optimized code version is left
and one compiled without inlining and other optimizations is entered.
Self uses scope descriptors to describe the frames
that need to be re-created when leaving the optimized code.
The scope descriptors are between 0.42 and 1.09 times
the size of the generated machine code.
The information needed for debugging together
is between 1.22 and 2.33 times the size of generated machine code,
according to the paper.

@@

Additionally we presented experimental data collected using the standard PyPy
benchmark set to evaluate previous observations and assumptions about guards. Our
experiments confirmed that guards are a very common
operation in traces. At the same time guards are associated with a high
overhead, because for all compiled guards information needs to be
stored to restore the execution state in case of a bailout. The measurements
showed that the compression techniques used in PyPy effectively reduce the
overhead of guards, but they still produce a significant overhead. The results
also showed that guard failure is a local event: there are few
guards that fail at all, and even fewer that fail very often.
These numbers validate the design decision of reducing the overhead of
successful guard checks as much as possible while paying a higher price in the
case of bailout due to having to decode a compressed state representation.
The compressed state representation reduces the memory footprint of rarely
used data.
